---
title: "Structuring around exemplars - ESIDA workshop"
author: "Joris Falip"
date: "12/2018"
output: html_document
params:
   dataset: "residential"
   k: 124
   threshold: 10
   range: 50
   nb_cpus: 2
editor_options: 
  chunk_output_type: console
---

# Setup

This section highlights how to install the needed packages using two methods, then download et read the datasets used

### Packages

You can install every needed package by hand, or run one of the two following chunks. Those two chunks won't run by default, you need to change _eval_ to _TRUE_.

##### Checkpoint

This package is the best solution for reproducibility. It will install every needed package in the same version that was available on the day the analysis was done. This process takes a long time.
Before running this chunk, you need to install the _checkpoint_ package then create a _.checkpoint_ folder inside the folder containing this Rmarkdown file.

```{r packages-checkpoint, eval=FALSE}
library(checkpoint)
checkpoint("2018-12-17")
```

##### Automagic

This package installs (using _install.packages()_) the missing packages, avoiding the hassle of installing packages one by one.
Before running this chunk, you need to install the _automagic_ package and set your current working directory to the folder containing this Rmd file.

```{r packages-automagic, eval=FALSE}
library(automagic)
#automagic()
```

We can now load the required packages and the proposed structuration algorithm.

```{r setup}
library(fields)
library(foreach)
library(doParallel)
library(magrittr)
library(ggpubr)
library(ggplot2)
library(readxl)
library(readr)
library(purrr)
library(igraph)
library(igraph)

source("structuration.R")
```

We then download the two UCI datasets if they are not already in the data folder.

```{r datasets-download}
if (!file.exists("./data/Residential-Building-Data-Set.xlsx")) {
  download.file("https://archive.ics.uci.edu/ml/machine-learning-databases/00437/Residential-Building-Data-Set.xlsx", "./data/Residential-Building-Data-Set.xlsx")
}

if (!file.exists("./data/CommViolPredUnnormalizedData.txt")) {
  download.file("http://archive.ics.uci.edu/ml/machine-learning-databases/00211/CommViolPredUnnormalizedData.txt", "./data/CommViolPredUnnormalizedData.txt")
}
```

We load the two UCI datasets and generate the synthethic dataset.

```{r load-studied-data}
if (params$dataset == "residential") {
  ### RESIDENTIAL ###
  # https://archive.ics.uci.edu/ml/datasets/Residential+Building+Data+Set
  
  housing <- read_excel("data/Residential-Building-Data-Set.xlsx", 
                        sheet = "Data", col_names = FALSE, col_types = c("blank", 
                                                                         "blank", "blank", "blank", "numeric", 
                                                                         "numeric", "numeric", "numeric", 
                                                                         "numeric", "numeric", "numeric", 
                                                                         "numeric", "numeric", "numeric", 
                                                                         "numeric", "numeric", "numeric", 
                                                                         "numeric", "numeric", "numeric", 
                                                                         "numeric", "numeric", "numeric", 
                                                                         "numeric", "numeric", "numeric", 
                                                                         "numeric", "numeric", "numeric", 
                                                                         "numeric", "numeric", "numeric", 
                                                                         "numeric", "numeric", "numeric", 
                                                                         "numeric", "numeric", "numeric", 
                                                                         "numeric", "numeric", "numeric", 
                                                                         "numeric", "numeric", "numeric", 
                                                                         "numeric", "numeric", "numeric", 
                                                                         "numeric", "numeric", "numeric", 
                                                                         "numeric", "numeric", "numeric", 
                                                                         "numeric", "numeric", "numeric", 
                                                                         "numeric", "numeric", "numeric", 
                                                                         "numeric", "numeric", "numeric", 
                                                                         "numeric", "numeric", "numeric", 
                                                                         "numeric", "numeric", "numeric", 
                                                                         "numeric", "numeric", "numeric", 
                                                                         "numeric", "numeric", "numeric", 
                                                                         "numeric", "numeric", "numeric", 
                                                                         "numeric", "numeric", "numeric", 
                                                                         "numeric", "numeric", "numeric", 
                                                                         "numeric", "numeric", "numeric", 
                                                                         "numeric", "numeric", "numeric", 
                                                                         "numeric", "numeric", "numeric", 
                                                                         "numeric", "numeric", "numeric", 
                                                                         "numeric", "numeric", "numeric", 
                                                                         "numeric", "numeric", "numeric", 
                                                                         "numeric", "numeric", "numeric", 
                                                                         "numeric", "numeric", "numeric", 
                                                                         "numeric", "numeric"), skip = 2)
  
  data <- as.matrix(housing)
  soluce <- data[ ,105]
  data <- data[ ,-105]
  data <- data[ ,-104]
  
  rm(housing)
}

if (params$dataset == "communities") {
  ### CRIME ###
  # http://archive.ics.uci.edu/ml/datasets/Communities+and+Crime+Unnormalized
  
  crime <- read_csv("data/CommViolPredUnnormalizedData.txt", col_names = FALSE, na = "?")
  
  data <- crime[ ,-c(1:5)]
  data <- as.matrix(data)
  soluce <- crime$X131
  data <- data[ ,1:124]
  colToKeep <- colSums(is.na(data)) > 0
  data <- data[ ,!colToKeep]
  
  rm(crime,colToKeep)
}

if (params$dataset == "norm") {
  ### SYNTHETHIC ###
  
  data <- matrix(0, ncol = 1000, nrow = 300)
  set.seed(2019)
  for (norm in 1:800) {
    data[, norm] <- rnorm(300, mean = 0, sd = 1)
  }
  for (unif in 801:1000) {
    data[, unif] <- runif(300, 0, 1)
  }
}
```

# Analysis

To start the analysis, the _dataset_ and _k_ parameters are required, in the header of this Rmd file.
_dataset_ can take the following values : "residential" for the Residential Building dataset, "communities" for the Communities and Crime dataset, and "norm" for the synthethic dataset using normal distributions and uniform distributions.
_threshold_ parameter of this script is the threshold _T_ for formula 1 of the article.

The following chunk run the proposed exemplar and structures elements around exemplars. It then displays the structure as a graph, and compute the number of connected components, their size and mean diameter.

```{r prediction-joris}
results_exemplars <- structuration(data, params$k, params$threshold)
exemplars_graph <- make_graph(as.vector(rbind(1:length(results_exemplars), results_exemplars)))
plot(exemplars_graph, vertex.size = 3, edge.arrow.size = 0.1, vertex.label = "")
components(exemplars_graph)$no
mean(components(exemplars_graph)$csize)
mean(unlist(lapply(decompose(exemplars_graph), diameter)))
```

We create a similar structure by linking each element ot its closest neighbor using a minkowski distance of order 0.75. We extract the same information regarding connected components.

```{r nearest-minkowski}
distance_matrix <- as.matrix(dist(scale(data), method = "minkowski", p = 0.75))
results_minkowski <- integer(dim(data)[1])
diag(distance_matrix) <- 9999999
for (elem in 1:dim(data)[1]) {
  results_minkowski[elem] <- which(distance_matrix[elem,] == min(distance_matrix[elem,]))
}

mink_graph <- make_graph(as.vector(rbind(1:dim(data)[1], results_minkowski)))
plot(mink_graph, vertex.size = 3, edge.arrow.size = 0.1, vertex.label = "")
components(mink_graph)$no
mean(components(mink_graph)$csize)
mean(unlist(lapply(decompose(mink_graph), diameter)))
```

This chunk generates a chart displaying the number of connected components of the resulting structure, for every value of the input paramter _k_ ranging from 1 to input parameter _range_. This code is executed in a multithreaded fashion, using as many cores as specified by parameter *nb_cpus*.

```{r evolution-composantes}
registerDoParallel(params$nb_cpus)

composantes <- foreach(i = 1:params$range) %dopar% {
  results_components <- structuration(data, i, params$threshold)
  
  resgraph <- make_graph(as.vector(rbind(seq_len(dim(data)[1]), results_components)))
  return(components(resgraph)$no)
}

composantes <- unlist(composantes)
df <- as.data.frame(composantes)

composantes_graphs <- ggplot(df, aes(1:params$range)) + 
  geom_line(aes(y = composantes, color = "")) +
  labs(x = "Input parameter K", y = "Number of connected components")
composantes_graphs + theme_pubr() + grids(linetype = "dashed") + theme(legend.position = "none")

stopImplicitCluster()
```